---
title: "Practical Machine Learning - HAR Project"
author: "Sam S. Kim"
date: "April 23, 2015"
output: html_document
---

### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible 
to collect a large amount of data about personal activity relatively inexpensively. 
These type of devices are part of the quantified self movement – a group of 
enthusiasts who take measurements about themselves regularly to improve their 
health, to find patterns in their behavior, or because they are tech geeks. One 
thing that people regularly do is quantify how much of a particular activity they 
do, but they rarely quantify how well they do it.

**In this project,** I'll be using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.  They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website and it's cited at the bottom of this document.


#### Explore and Prepare Data

Load the given data sets:

```{r results='hide'}
training <- read.csv("data/pml-training.csv")
testing <- read.csv("data/pml-testing.csv")
```

Then review the summary:

```{r results='hide'}
# Briefly review the summary of the dataset - each variable
str(training)
summary(training)

str(testing)
summary(testing)
```

#### Two things to note from the summary:

**One.** The response variable *classe* has 5 levels:

A = Exactly according to the specification

B = Throwing the elbows to the front

C = Lifting the dumbbell only halfway

D = Lowering the dumbbell only halfway

E = Throwing the hips to the front

Also note the proportion of the *classe* value distributions:

```{r echo=FALSE}
prop.table(table(training$classe))
```


**Two.** There were lots of missing data, or NA's.

Since I'd need to predict the *classe* for each observation in testing set, 
I'd to make sure the training and testing sets have the same set of predictors, and 
they could not have missing values.  So I removed all the columns in the testing set 
with all NA's.

```{r}
# Get column names that have zero NA's.  This is the possible pool of predictors we could use.
colsForTesting <- colnames(testing)[colSums(is.na(testing)) == 0]
length(colsForTesting)

# Now get the column names for training set;  Include the "classe" response variable for training.
colsForTraining <- c(colsForTesting[1:59], "classe")
length(colsForTraining)
```


### Processing & Modeling

Load the necessary libraries:
```{r results='hide', message=FALSE}
library(caTools)
library(randomForest)
library(caret)

library(doMC)
registerDoMC(cores = 3)  # Use multi cores for faster processing of random forest.
```


Subset to just the necessary columns and split the training data to train (70%) & validation (30%) data.

```{r results='hide'}
# Subset the only the necessary columns
training <- training[, colsForTraining]
training$X <- NULL

# Split the data into train & test set
set.seed(300)
spl <- sample.split(training$classe, SplitRatio=.70)
mTrain <- subset(training, spl == TRUE)
mTest <- subset(training, spl == FALSE)

``` 


Apply Random Forest using caret package, and apply k-fold cross valiation.

```{r cache=TRUE}
# K-fold cross validation
fitControl <- trainControl(method="cv", number=10)

set.seed(300)
modFitRF <- train(classe ~., data=mTrain, method="rf", trControl=fitControl)
```

### Out-of-sample error

Now predicted the *classe* on the validation set splitted from the training data.
As shown below in the confusion matrix, that accuracy is 0.9992.
That's an out-of-sample error of just **0.08%**.

```{r}
predRF <- predict(modFitRF, mTest)

confusionMatrix(predRF, mTest$classe)
```


### Predict the *classe* in the testing set using the model just built.

```{r}
predTesting <- predict(modFitRF, testing)

predTesting
```



### Conclusion

So a random forest model did the trick of guessing *classe* correctly for the testing set.
Many of the measurements from accelerometers turned out to be very good predictors for how 
well a certain activity is performed.  Perhaps it could lead to a 
self-teaching of physical activity with the aid of a machine that would give immediate 
feedback based on these similar sets of data.



#### Reference:
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. [Qualitative Activity Recognition 
of Weight Lifting Exercises](http://groupware.les.inf.puc-rio.br/work.jsf?p1=11201). Proceedings of 4th International Conference in Cooperation with SIGCHI 
(Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
 
[DocumentoDocumento](http://groupware.les.inf.puc-rio.br/public/papers/2013.Velloso.QAR-WLE.pdf)




